---
title: "BRCA_evidence_status"
author: "Weston Bodily"
date: "December 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

```{r}
library(readr)
library(Rtsne)
library(dplyr)
library(ggplot2)
library(reshape)
library(extremevalues)
```

## Import the necessary datasets
##    NOTE: It's problematic to generate the Mutational Signature information anew, 
##    so the "BRCA-like" information must be obtained by first running the script named
##    "Rtsine_mut_sigs_new.Rmd" to get the "brca-likeness_mut_sig.txt" file first.
##    Other data sets have BRCA-likeness determined in this script.

```{r}
methylationData <- as.data.frame(read_tsv("../Data/methyl_brca_re.tsv"))

germlineData <- as.data.frame(read_tsv("../Data/germline_snv_indel.tsv"))
somaticData <- as.data.frame(read_tsv("../Data/Somatic_ver_1.19.17.txt"))

CNVData <- as.data.frame(read_tsv("../Data/somatic_CNV.tsv"))

TCGASubtypeData <- (read_tsv("../Data/TCGA_Subtypes_Mutations_BasicTabFormatted.txt"))
TCGASubtypeData <- TCGASubtypeData[,colSums(is.na(TCGASubtypeData))<nrow(TCGASubtypeData)]

```

## Preparatory data formatting

```{r}
uniqueSomaticGenes <- sort(unique(somaticData$Gene))
uniqueSomaticSamples <- sort(unique(somaticData$Sample)) ##Added

BRCAGenes <- c("BRCA1", "BRCA2")
```

## CNV Data - generates insertionCNV, deletionCNV, deletionCNV Dataframes

```{r}
somaticGene_Symbol <- table(CNVData$Gene_Symbol)
somaticGenes <- c("BRCA1", "BRCA2") #, "RAD51D", "RAD52", "RAD54B", "NBN",  "MUS81", "EME2", "GIYD1", "GIYD2")
somaticPlotData <- filter(CNVData, Gene_Symbol %in% somaticGenes)
somaticPlotData$Gene_Symbol <- factor(somaticPlotData$Gene_Symbol)

CNVFilterFunction <- function(CopyValue) {

  tempNumbers <- c(CopyValue)
  subtypeSamples <- sort(unique(TCGASubtypeData$Patient_ID))
  subtypeCategory <- unique(sort(TCGASubtypeData$PAM50))
  subtypeData <- filter(TCGASubtypeData, PAM50 %in% subtypeCategory)
  CNVPlotData <- filter(somaticPlotData, Value %in% tempNumbers)

  subtypeDataFrame <- NULL

  CNVPlotData$Sample = substr(CNVPlotData$Sample,1,nchar(CNVPlotData$Sample)-3)

  comboContainer <- inner_join(select(subtypeData, Patient_ID, PAM50), CNVPlotData, by = c("Patient_ID" = "Sample"))
  comboContainer$PAM50 <- as.factor(comboContainer$PAM50)
  
  return(comboContainer)

}

insertionCNV <- CNVFilterFunction(2)
deletionCNV <- CNVFilterFunction(-2)
tempNum <- c(2, -2)
combinedCNV <- CNVFilterFunction(tempNum)

```

# Methylation Data setup (note to self: remember to use germlineData instead of germlineMutationData)

```{r}
HRR_genes <- c("BRCA1", "BRCA2", "RAD51", "RAD51B", "RAD51D", "DMC1", "XRCC2", "XRCC3", "RAD52", "RAD54L", "RAD54B", "SHFM1", "RAD50", "MRE11A", "NBN", "RBBP8", "MUS81", "EME1", "EME2", "GIYD1", "GIYD2", "GEN1")

germlineMutationFrequencies <- table(germlineData$Gene)
topGermlineMutationGenes <- names(head(sort(germlineMutationFrequencies, decreasing=TRUE), n=8))

methylationProbe <- table(methylationData$Probe)
methylationGenes <- topGermlineMutationGenes
methylationPlotData <- filter(methylationData, Probe %in% methylationGenes)
methylationPlotData$Probe <- factor(methylationPlotData$Probe)
uniqueMethylationSamples <- sort(unique(methylationData$Sample))

```

# Methylation data continued, generates "outlierContainer"

```{r}
geneLoopList <- c("BRCA1", "BRCA2", "RAD54L", "SHFM1", "RBBP8")

limitContainer <- NULL
outlierContainer <- NULL

for (gene in geneLoopList)
{
  geneContainer <- gene
  methylationGene <- as.data.frame(filter(methylationData, Probe %in% geneContainer))
  methylationSamples <- methylationGene$Sample
  methylationValues <- methylationGene$Value

  if(gene %in% c("BRCA1"))
  {
    distribution <- "exponential"
  }
  else
  {
    distribution <- "exponential"
  }

  methylationOutliers <- getOutliersI(methylationValues, rho=c(1, 1), FLim=c(0.1, 0.9), distribution=distribution)
  outlierPlot(methylationGene$Value, methylationOutliers, mode="qq")
  outlierSamples <- methylationSamples[methylationOutliers$iRight]

  for (sample in outlierSamples)
  {
    tempCont <- c(sample, gene)
    outlierContainer <- rbind(outlierContainer, tempCont)
  }
  colnames(outlierContainer) <- c("Sample", "Gene")
}

outlierContainer <- as.data.frame(outlierContainer)
```

# Iterate through data sets and detect a sample by TCGA_ID, determine BRCA-like status, save info to DF
# DF's to look through:
#   outlierContainer (methylation)
#   deletionCNV (CNV)
#   mutSigData (Mutational signatures)
#   somaticData (somatic mutations)
#   germlineData (germline mutations)
# Saves to a DF called "cMat"

```{r}
  cMat <- NULL
  for (sample_s in uniqueSomaticSamples)
  {
    BRCAstatus <- NULL
    gene_one <- "BRCA1"
    gene_two <- "BRCA2"
    mut_sig_name <- "Mutation_sig"
    
    geneCounter <- 0
    geneSomaticB1 <- filter(somaticData, Gene==gene_one&Sample==sample_s)
    geneGermlineB1 <- filter(germlineData, Gene==gene_one&Sample_ID==sample_s)
    geneSomaticB2 <- filter(somaticData, Gene==gene_two&Sample==sample_s)
    geneGermlineB2 <- filter(germlineData, Gene==gene_two&Sample_ID==sample_s)
    
    likeMethyl <- filter(outlierContainer, Sample==sample_s)
    likeCNV <- filter(deletionCNV, Patient_ID==sample_s)

    geneTest1 <- filter(somaticData, Sample==sample_s)
    geneTest2 <- filter(germlineData, Sample_ID==sample_s)
    geneSampleValue <- ""
    
    if ( nrow(geneGermlineB1) > 0) 
      {
        if(nrow(geneGermlineB2) > 0)
        {       geneSampleValue <- paste("BRCA1_and_BRCA2", sep="")      }
        else  
        {       geneSampleValue <- paste("BRCA1", sep="")     }
      }
    else if (nrow(geneGermlineB2) > 0)
      {     geneSampleValue <- paste("BRCA2", sep="")     }
    else if (nrow(geneSomaticB1) > 0 || nrow(geneSomaticB2) > 0)
      {     geneSampleValue <- paste("BRCA-like", sep="")      }
    else if (nrow(likeMethyl) > 0 || nrow(likeCNV) > 0 ) # || nrow(likeMutSig) > 0 )
      {
        geneSampleValue <- paste("BRCA-like", sep="")
      }
    else {  geneSampleValue <- paste("Other", sep="")  }
    
    if (geneSampleValue != "No_data")
      { BRCAstatus <- c(sample_s, geneSampleValue) }
  
  if (is.null(BRCAstatus) == FALSE)
    { cMat <- rbind(cMat, BRCAstatus) }
  }
  colnames(cMat) <- c("Sample", "BRCA_status")

```

# BRCA Evidence

```{r}
  eMat <- NULL
  for (sample_s in uniqueSomaticSamples)
  {
    BRCAstatus <- NULL
    gene_one <- "BRCA1"
    gene_two <- "BRCA2"
    mut_sig_name <- "Mutation_sig"
    
    geneCounter <- 0
    geneSomaticB1 <- filter(somaticData, Gene==gene_one&Sample==sample_s)
    geneGermlineB1 <- filter(germlineData, Gene==gene_one&Sample_ID==sample_s)
    geneSomaticB2 <- filter(somaticData, Gene==gene_two&Sample==sample_s)
    geneGermlineB2 <- filter(germlineData, Gene==gene_two&Sample_ID==sample_s)
    
    likeMethylB1 <- filter(outlierContainer, Sample==sample_s&Gene=="BRCA1")
    likeMethylB2 <- filter(outlierContainer, Sample==sample_s&Gene=="BRCA2")
    likeMethyl_O <- filter(outlierContainer, Sample==sample_s)
    likeCNV1 <- filter(deletionCNV, Patient_ID==sample_s&Gene_Symbol=="BRCA1")
    likeCNV2 <- filter(deletionCNV, Patient_ID==sample_s&Gene_Symbol=="BRCA2")

    geneTest1 <- filter(somaticData, Sample==sample_s)
    geneTest2 <- filter(germlineData, Sample_ID==sample_s)
    geneSampleValue <- ""
    
    if (nrow(geneGermlineB1) > 0 & nrow(geneGermlineB2) > 0)
    {     
      geneSampleValue <- paste("Germline_BRCA1&2_mutation", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    else if (nrow(geneGermlineB1) > 0 ) 
    {     
      geneSampleValue <- paste("Germline_BRCA1_mutation", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    else if (nrow(geneGermlineB2) > 0 ) 
    {     
      geneSampleValue <- paste("Germline_BRCA2_mutation", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    if (nrow(geneSomaticB1) > 0)
    {     
      geneSampleValue <- paste("Somatic_BRCA1_mutation", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    if (nrow(geneSomaticB2) > 0)
    {     
      geneSampleValue <- paste("Somatic_BRCA2_mutation", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    if (nrow(likeMethylB1) > 0 )
    {     
      geneSampleValue <- paste("Methylation_BRCA1", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    else if (nrow(likeMethylB2) > 0 )
    {     
      geneSampleValue <- paste("Methylation_BRCA2", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
        }
    else if (nrow(likeMethyl_O) > 0 )
    {     
      geneSampleValue <- paste("Methylation_Other", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    if (nrow(likeCNV1) > 0 )
    {     
      geneSampleValue <- paste("Deletion_BRCA1", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    if (nrow(likeCNV2) > 0 )
    {     
      geneSampleValue <- paste("Deletion_BRCA2", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }

    if (geneSampleValue == "")
    {
      geneSampleValue <- paste("None", sep="")      
      BRCAstatus <- c(sample_s, geneSampleValue)
      eMat <- rbind(eMat, BRCAstatus)
    }
    
  }
  colnames(eMat) <- c("Sample", "BRCA_evidence")

```

# Save DFs to files

```{r}
write.table(cMat, file="BRCA-like_status.txt", append= FALSE, quote=FALSE, sep="\t", row.names=FALSE)
write.table(eMat, file="BRCA-like_evidence.txt", append= FALSE, quote=FALSE, sep="\t", row.names=FALSE)
```
