---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

```{r}
library(readr)
library(Rtsne)
library(dplyr)
library(ggplot2)
library(reshape)
```

## HiSeq data importing

```{r}
hiseq_data <- (read_tsv("../Data/brca_hiseq.txt"))
hiseq_IDs <- as.vector(as.data.frame(hiseq_data)[,1])
```

## PAM50 genes
##  Altered, because one gene wasn't present in the available data

```{r}
pam_gene_altered <- c("UBE2T", "BIRC5", "NUF2", "CDC6", "CCNB1", "TYMS", "MYBL2", "CEP55", "MELK", "NDC80", "RRM2", "UBE2C", "CENPF", "PTTG1", "EXO1", "ANLN", "CCNE1", "CDC20", "MKI67", "KIF2C", "ACTR3B", "MYC", "EGFR", "KRT5", "PHGDH", "CDH3", "MIA", "KRT17", "FOXC1", "SFRP1", "KRT14", "ESR1", "SLC39A6", "BAG1", "MAPT", "PGR", "CXXC5", "MLPH", "BCL2", "MDM2", "NAT1", "FOXA1", "BLVRA", "MMP11", "GPR160", "FGFR4", "GRB7", "TMEM45B", "ERBB2")

gene_pam <- as.matrix(as.data.frame(subset(hiseq_data, select=pam_gene_altered)))
row.names(gene_pam) <- hiseq_IDs
```

## Remove non-tumor (01) samples

```{r}
samplesToKeep <- NULL

for (sampleID in rownames(gene_pam))
{
  tissueType <- substring(strsplit(sampleID, "-")[[1]][4], 1, 2)
  if (tissueType == "01")
    samplesToKeep <- c(samplesToKeep, sampleID)
}

gene_pam <- gene_pam[samplesToKeep,]
```

## Merge replicate samples

```{r}
shortenSampleIDs <- function(longSampleIDs)
{  return(as.character(substr(longSampleIDs,1,12)))  }

shortSampleIDs <- shortenSampleIDs(rownames(gene_pam))
uniqueShortSampleIDs <- unique(shortSampleIDs)
mod_gene_pam <- NULL

for (sampleID in uniqueShortSampleIDs)
{
  numReplicates <- table(shortSampleIDs)[sampleID]
  if (numReplicates == 1) {
    sampleData <- gene_pam[grep(sampleID, rownames(gene_pam)),,drop=FALSE]
  } else {
    sampleData <- gene_pam[which(shortenSampleIDs(rownames(gene_pam)) == sampleID),]
    sampleData <- matrix(apply(sampleData, 2, median), nrow=1)
    colnames(sampleData) <- colnames(gene_pam)
  }

  rownames(sampleData) <- sampleID  
  mod_gene_pam <- rbind(mod_gene_pam, sampleData)
}

print(dim(mod_gene_pam))
```

## Run Rtsne on the imported data set, and select the output values for a DF

```{r}
modRowNames <- row.names(mod_gene_pam)
log_gene_pam <- log2(mod_gene_pam + 1)

set.seed(0)
rtsne_out <- Rtsne(log_gene_pam, check_duplicates = FALSE, verbose = TRUE)
plot(rtsne_out$Y)

rtsne_y <- rtsne_out$Y
row.names(rtsne_y) <- modRowNames
```

## This section is for getting BRCA1/2 data

```{r}
germlineData <- as.data.frame(read_tsv("../Data/germline_snv_indel.tsv"))
somaticData <- as.data.frame(read_tsv("../Data/Somatic_ver_1.19.17.txt"))
```

## Create data matrix for samples with BRCA1/2 status
## This will exclude samples if not found in the germline of somatic data files

```{r}
cMat <- NULL
for (sample_s in modRowNames)
{
  BRCAstatus <- NULL
  gene_one <- "BRCA1"
  gene_two <- "BRCA2"
  {
    geneCounter <- 0
    #geneSomaticB1 <- filter(somaticData, Gene==gene_one&Sample==sample_s)
    geneGermlineB1 <- filter(germlineData, Gene==gene_one&Sample_ID==sample_s)
    #geneSomaticB2 <- filter(somaticData, Gene==gene_two&Sample==sample_s)
    geneGermlineB2 <- filter(germlineData, Gene==gene_two&Sample_ID==sample_s)
    geneTest1 <- filter(somaticData, Sample==sample_s)
    geneTest2 <- filter(germlineData, Sample_ID==sample_s)
    geneSampleValue <- ""
    
    if (nrow(geneGermlineB1) > 0) 
      {
        if(nrow(geneGermlineB2) > 0)
        {
          geneSampleValue <- paste("BRCA1 and BRCA2", sep="") 
        }
        else  
        {
          geneSampleValue <- paste("BRCA1", sep="")
        }
      }
    else if (nrow(geneGermlineB2) > 0)
      {
        geneSampleValue <- paste("BRCA2", sep="")
      }
    else if (nrow(geneTest1) > 0 )
      { geneSampleValue <- paste("Other", sep="") }  
    else 
      { geneSampleValue <- paste("No_data", sep="") }
   
    if (geneSampleValue != "No_data")
      { BRCAstatus <- c(sample_s, geneSampleValue) }
  }
  if (is.null(BRCAstatus) == FALSE)
    { cMat <- rbind(cMat, BRCAstatus) }
}

colnames(cMat) <- c("Sample", "BRCA_status")
```

## data setup

```{r}
gMat <- as.data.frame(cMat)
gMat <- unique(gMat)

rtsne_citrus <- NULL
rtsne_row_names <- row.names(rtsne_y)
```

```{r}
#alist<-c(1:1095)
alist<-c(1:1097)
zMat <- NULL
z_element <- NULL

apply(gMat, 1, function(row)
  {
    sample_id <- NULL
    status <- NULL
    num_x <- 0
    num_y <- 0
    z_element <- NULL
    
    for (item in alist) 
      {
        if (row[1] == rtsne_row_names[item]) 
        { 
          num_x <- rtsne_y[item,1]
          num_y <- rtsne_y[item,2]
        } 
      }  
    sample_id <- as.character(row[1])
    status <- as.character(row[2])
    z_element <- c(sample_id, status, num_x, num_y)
    zMat <<- rbind(zMat, z_element)
    } 
  ) 

z_col_names <- c("Sample_ID", "BRCA_status", "Val_1", "Val_2")
colnames(zMat) <- z_col_names

head(zMat)
```

## Get the subtype data for subtype plotting, save it to rtsne_citrus

```{r}
TCGASubtypeData <- (read_tsv("../Data/TCGA_Subtypes_Mutations_BasicTabFormatted.txt"))
TCGASubtypeData <- TCGASubtypeData[,colSums(is.na(TCGASubtypeData))<nrow(TCGASubtypeData)]

subtypeSamples <- sort(unique(TCGASubtypeData$Patient_ID))
subtypeCategory <- unique(sort(TCGASubtypeData$PAM50))
subtypeData <- filter(TCGASubtypeData, PAM50 %in% subtypeCategory)

subtypeShort <- data.frame(matrix(NA, nrow = 1002, ncol = 2))
subtypeShort[,1] <- subtypeData[,1]
subtypeShort[,2] <- subtypeData[,10]

colnames(subtypeShort) <- c("Sample_ID", "PAM50")

rtsne_citrus <- inner_join(as.data.frame(zMat), as.data.frame(subtypeShort), by="Sample_ID" )
```

## Get BRCA status from previously saved file

```{r}
brcaEvidenceData <- (read_tsv("../Data/BRCA-like_evidence.txt"))
colnames(brcaEvidenceData) <- c("Sample_ID", "BRCA_evidence")
rtsne_evidence <- inner_join(as.data.frame(zMat), as.data.frame(brcaEvidenceData), by="Sample_ID" )
```

## Formatting for ease of methylation status plotting

```{r}
met_stat <- NULL
new_column <- sapply(rtsne_evidence$BRCA_evidence, function(x)
  {
    met_stat <- ""
    if(x == "Methylation_BRCA1") met_stat = "M1"
    else if (x == "Methylation_BRCA2") met_stat = "M2"
    else if (x == "Methylation_Other") met_stat = "MO"
    else { met_stat = "N" }
    return(met_stat)
  })

rtsne_evidence$Met <- new_column
```

## Formatting for ease of CNV deletion status plotting

```{r}
del_stat <- NULL
new_column <- sapply(rtsne_evidence$BRCA_evidence, function(x)
  {
    del_stat <- ""
    if(x == "Deletion_BRCA1") del_stat = "D1"
    else if (x == "Deletion_BRCA2") del_stat = "D2"
    else if (x == "Deletion_both") del_stat = "D9"
    else { del_stat = "N" }
    return(del_stat)
  })

rtsne_evidence$CNV <- new_column
```

## Formatting for ease of Somatic mutation plotting

```{r}
Som_stat <- NULL
new_column <- sapply(rtsne_evidence$BRCA_evidence, function(x)
  {
    Som_stat <- ""
    if(x == "Somatic_BRCA1_mutation") Som_stat = "S1"
    else if (x == "Somatic_BRCA2_mutation") Som_stat = "S2"
    else if (x == "Somatic_BRCA1&2_mutation") Som_stat = "S9"
    else { Som_stat = "N" }
    return(Som_stat)
  })

rtsne_evidence$Som <- new_column
```

## Create new columns for "BRCA1" with the type of aberration

```{r}
Som_stat <- NULL
new_column <- sapply(rtsne_evidence$BRCA_evidence, function(x)
  {
    Som_stat <- ""
    if(x == "Somatic_BRCA1_mutation") Som_stat = "S"
    else if (x == "Germline_BRCA1_mutation") Som_stat = "G"
    else if (x == "Deletion_BRCA1") Som_stat = "D"
    else if (x == "Methylation_BRCA1") Som_stat = "M"
    else { Som_stat = "N" }
    return(Som_stat)
  })

rtsne_evidence$B1A <- new_column

Som_stat <- NULL
new_column <- sapply(rtsne_evidence$BRCA_evidence, function(x)
  {
    Som_stat <- ""
    if(x == "Somatic_BRCA1_mutation") Som_stat = "Y"
    else if (x == "Germline_BRCA1_mutation") Som_stat = "Y"
    else if (x == "Deletion_BRCA1") Som_stat = "Y"
    else if (x == "Methylation_BRCA1") Som_stat = "Y"
    else { Som_stat = "N" }
    return(Som_stat)
  })

rtsne_evidence$AnyBRCA1 <- new_column
```

## Create new columns for "BRCA2" with the type of aberration

```{r}
Som_stat <- NULL
new_column <- sapply(rtsne_evidence$BRCA_evidence, function(x)
  {
    Som_stat <- ""
    if(x == "Somatic_BRCA2_mutation") Som_stat = "S"
    else if (x == "Germline_BRCA2_mutation") Som_stat = "G"
    else if (x == "Deletion_BRCA2") Som_stat = "D"
    else if (x == "Methylation_BRCA2") Som_stat = "M"
    else { Som_stat = "N" }
    return(Som_stat)
  })

rtsne_evidence$B2A <- new_column

Som_stat <- NULL
new_column <- sapply(rtsne_evidence$BRCA_evidence, function(x)
  {
    Som_stat <- ""
    if(x == "Somatic_BRCA2_mutation") Som_stat = "Y"
    else if (x == "Germline_BRCA2_mutation") Som_stat = "Y"
    else if (x == "Deletion_BRCA2") Som_stat = "Y"
    else if (x == "Methylation_BRCA2") Som_stat = "Y"
    else { Som_stat = "N" }
    return(Som_stat)
  })

rtsne_evidence$AnyBRCA2 <- new_column
```

## Save the rtsne_evidence DF to a file for use in later analyses.

```{r}
write.table(rtsne_evidence, file="rtsne_evidence_DF.txt", append= FALSE, quote=FALSE, sep="\t", row.names=FALSE)
```