---
title: "Heat Map"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NOTE: this file was used to generate "waterfall" plots, which I still have. 
##  Unfortunately, ComplexHeatmap isn't supported by current versions of R,
##  so I can't test to see if this script still works for now
## Prep

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
```

## Read germline/somatic mutation data

```{r}
germlineData <- as.data.frame(read_tsv("../Data/germline_snv_indel.tsv"))
somaticData <- as.data.frame(read_tsv("../Data/BRCA_parsed_hg38.txt"))
```

## Germline - Build matrices
```{r}
uniqueGermlineGenes <- sort(unique(germlineData$Gene))
uniqueSomaticGenes <- sort(unique(somaticData$Gene))
uniqueSamples <- sort(unique(germlineData$Sample_ID))
uniqueSomaticSamples <- sort(unique(somaticData$Sample)) ##Added

uniqueTotalSamples <- c(uniqueSamples, uniqueSomaticSamples) ##A
uniqueTotalSamples <- sort(unique(uniqueTotalSamples$Sample_ID)) ##A
```

```{r}
test_lemon = 0
test_lemon = test_lemon + 1
print(test_lemon)
test_lemon = test_lemon + 2
print(test_lemon)
```

```{r}
BRCA1_g_count = 0
BRCA2_g_count = 0

gMat <- NULL
for (gene in uniqueGermlineGenes)
{
  geneValues <- NULL
  for (sample in uniqueSamples)
  {
    geneSampleDataFrame <- filter(germlineData, Gene==gene&Sample_ID==sample)
    
    geneSampleValue <- ""
    if (nrow(geneSampleDataFrame) > 0)
      {
        if (gene=="BRCA1") { BRCA1_g_count <- BRCA1_g_count + 1 }
        if (gene=="BRCA2") { BRCA2_g_count <- BRCA2_g_count + 1 }
        geneSampleValue <- paste(geneSampleDataFrame$Mutation_Type, ";", sep="")
      }
      
    geneValues <- c(geneValues, geneSampleValue)
  }
  
  gMat <- rbind(gMat, geneValues)
}

rownames(gMat) <- uniqueGermlineGenes
colnames(gMat) <- uniqueSamples
```

## Germline - Create heatmap

```{r}
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#FFFFFF", col = NA))
    },
    Insertion = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#206490", col = NA))
    },
    SNV = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#35ABBC", col = NA))
    },
    Deletion = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#4AE48B", col = NA))
    }
)

col = c("Insertion" = "#206490", "SNV" = "#35ABBC","Deletion" = "#4AE48B")

mutationOptions <- c("Insertion", "SNV", "Deletion")

oncoPrint(gMat, get_type = function(x) strsplit(x, ";")[[1]],
    alter_fun = alter_fun, col = col, 
    column_title = "Germline Mutations",
    heatmap_legend_param = list(title = "Value", at = mutationOptions,
        labels = mutationOptions))
```

## Somatic data - build matricies

```{r}
##uniqueSomaticGenes <- sort(unique(somaticData$Gene))
uniqueSomaticGenes <- uniqueGermlineGenes
uniqueSomaticSamples <- NULL
uniqueSomaticSamples <- sort(unique(somaticData$Sample)) ##Added

geneCounter <- 0
BRCA1_s_count <- 0
BRCA2_s_count <- 0

sMat <- NULL
for (gene in uniqueSomaticGenes)
{
  geneValues <- NULL
  for (sample in uniqueSomaticSamples)
  {
    geneCounter <- 0
    geneSampleDataFrame <- filter(somaticData, Gene==gene&Sample==sample)
    
    geneSampleValue <- ""
    if (nrow(geneSampleDataFrame) > 1) 
      {
        for (row in geneSampleDataFrame) 
        {
          geneCounter <- (geneCounter + 1)
          if (gene == "BRCA1") { BRCA1_s_count <- BRCA1_s_count + 1 }
          if (gene == "BRCA2") { BRCA2_s_count <- BRCA2_s_count + 1 }
        }
        geneSampleValue <- paste("Multiple;", sep="")
      } 
    else if(nrow(geneSampleDataFrame) > 0) 
      {
        if (gene == "BRCA1") { BRCA1_s_count <- BRCA1_s_count + 1 }
        if (gene == "BRCA2") { BRCA2_s_count <- BRCA2_s_count + 1 }
        geneSampleValue <- paste(geneSampleDataFrame$Variant_Class, ";", sep="")
      }
    geneValues <- c(geneValues, geneSampleValue)
  }
  
  sMat <- rbind(sMat, geneValues)
}

rownames(sMat) <- uniqueSomaticGenes
colnames(sMat) <- uniqueSomaticSamples

print(BRCA1_s_count)
print(BRCA2_s_count)
```

## Somatic data - create heatmap

```{r}
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#FFFFFF", col = NA))
    },
    Multiple = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#6C14FE", col = NA))
    },
    insertion = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#D81738", col = NA))
    },
    SNV = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#E48E1E", col = NA))
    },
    deletion = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#4AE48B", col = NA))
    }

)

col = c("Multiple" = "#6C14FE", "insertion" = "#D81738", "SNV" = "#E48E1E","deletion" = "#4AE48B") 

mutationOptions <- c("Multiple", "insertion", "SNV", "deletion")

oncoPrint(sMat, get_type = function(x) strsplit(x, ";")[[1]],
    alter_fun = alter_fun, col = col, axis_gp = gpar(fontsize = 8),
    column_title = "Somatic Mutations",
    heatmap_legend_param = list(title = "Value", at = mutationOptions,
        labels = mutationOptions))
```

## Import Methylation data

```{r}
methylationData <- read_tsv("../Data/methyl_brca_re.tsv")

methylationProbe <- table(methylationData$Probe)
methylationGenes <- uniqueSomaticGenes

methylationPlotData <- filter(methylationData, Probe %in% methylationGenes)

methylationPlotData$Probe <- factor(methylationPlotData$Probe)
uniqueMethylationSamples <- sort(unique(methylationData$Sample))
```

## Filter samples (>= mean + 2 * stdv)

```{r}
containGenes <- uniqueSomaticGenes
methylatedSamples <- NULL

geneHyperCutoff <- NULL
cuttofDF <- NULL

for(gene in containGenes)
{

  methylationGene <- filter(methylationPlotData, Probe %in% gene)
  valueGene <- methylationGene$Value
  stdvGene <- sd(valueGene)
  meanGene <- mean(valueGene)
  hypCut <- (meanGene + (2 * stdvGene))
  if (!is.nan(hypCut))
  {
    cutoffVal <- c(gene, hypCut)
    cutoffDF <- rbind(cutoffDF, cutoffVal)
  }
  print(gene)
  print(hypCut)
  
  for(sample in valueGene)
  {
    outlierSamples <- NULL
    outlierValues <- NULL
    if (sample >= (meanGene + (2 * stdvGene)))
    {
      outlierValues <- sample
    }
    outlierSamples <- filter(methylationGene, Value %in% outlierValues)
    
    methylatedSamples <- rbind(methylatedSamples, outlierSamples)
  
  }
cutoffValue <- (meanGene + (2 * stdvGene))
tempCutoff <- c(gene, cutoffValue)
geneHyperCutoff <- rbind(geneHyperCutoff, tempCutoff)
}
```

## Write hypermethylation status to file for later use
## Note - this code currently isn't working, not sure what it's problem is but it doesn't want to cooperate.

```{r}

sampleMetContainer <- NULL
metSampleNames <- unique(methylationData$Sample)
metSampleProbes <- unique(methylationData$Probe)

for(sample in metSampleNames)
{
  for(probe in metSampleProbes)
  {
    tempMetContainer <- NULL
    lineHypContainer <- NULL
    lineMetContainer <- NULL
    lineHypContainer <- filter(methylatedSamples, Sample==sample&&Probe==probe)
    lineMetContainer <- filter(methylationData, Sample==sample&&Probe==probe)
    
    if (nrow(lineHypContainer) > 0 && nrow(lineMetContainer) > 0)
    {
      tempMetContainer <- c(lineMetContainer, "Hypermethylated")
    }
    else if (nrow(lineMetContainer) > 0)
    {
      tempMetContainer <- c(lineMetContainer, "Normal")
    }
    if(is.null(tempMetContainer) == FALSE)
    {
      sampleMetContainer <- rbind(sampleMetContainer, tempMetContainer)
    }
  }
}
```

## Combined data - build matricies

```{r}
combinedSamples <- uniqueSomaticSamples
uniqueCombinedSamples <- unique(combinedSamples)

geneCounter <- 0
```

## Create data matrix for combined data

```{r}
cMat <- NULL
for (gene in uniqueSomaticGenes)
{
  geneValues <- NULL
  for (sample in uniqueCombinedSamples)
  {
    geneCounter <- 0
    geneSomaticDataFrame <- filter(somaticData, Gene==gene&Sample==sample)
    geneGermlineDataFrame <- filter(germlineData, Gene==gene&Sample_ID==sample)
    geneSampleValue <- ""
  
    if (nrow(geneSomaticDataFrame) > 0) 
      {
        if(nrow(geneGermlineDataFrame) > 0)
        {
          geneSampleValue <- paste("Both;", sep="") 
        }
        else if(nrow(geneGermlineDataFrame) <= 0) 
        {
          geneSampleValue <- paste("Somatic;", sep="")
        }
      }
    else if (nrow(geneSomaticDataFrame) <= 0)
      {
        if(nrow(geneGermlineDataFrame) > 0)
        {
          geneSampleValue <- paste("Germline;", sep="") 
        }
      }
      
    geneValues <- c(geneValues, geneSampleValue)
  }
  cMat <- rbind(cMat, geneValues)
}

rownames(cMat) <- uniqueSomaticGenes
colnames(cMat) <- uniqueCombinedSamples
```

##Create heatmap for combined data

```{r}
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#FFFFFF", col = NA))
    },
    Germline = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#000099", col = NA))
    },
    Somatic = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#35ABBC", col = NA))
    },
    Both = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#e60000", col = NA))
    }

)

col = c("Germline" = "#000099", "Somatic" = "#35ABBC", "Both" = "#e60000") 

mutationOptions <- c("Germline", "Somatic", "Both")

oncoPrint(cMat, get_type = function(x) strsplit(x, ";")[[1]],
    alter_fun = alter_fun, col = col, 
    column_title = "Combined Mutations",
    heatmap_legend_param = list(title = "Value", at = mutationOptions,
        labels = mutationOptions))
```

## Create data matrix for combined data (including methylation data)

```{r}
cMat <- NULL
for (gene in uniqueSomaticGenes)
{
  geneValues <- NULL
  for (sample in uniqueCombinedSamples)
  {
    geneCounter <- 0
    geneSomaticDataFrame <- filter(somaticData, Gene==gene&Sample==sample)
    geneGermlineDataFrame <- filter(germlineData, Gene==gene&Sample_ID==sample)
    geneMethylationData <- filter(methylatedSamples, Probe==gene&Sample==sample)
    geneSampleValue <- ""
  
    if (nrow(geneSomaticDataFrame) > 0) 
      {
        if(nrow(geneGermlineDataFrame) > 0)
        {
          if(nrow(geneMethylationData) > 0 )  {  geneSampleValue <- paste("GerSomMet;", sep="")  }
          else                                {  geneSampleValue <- paste("GerSom;", sep="")  }
        }
        else if(nrow(geneGermlineDataFrame) <= 0) 
        {
          if(nrow(geneMethylationData) > 0 )  {  geneSampleValue <- paste("SomMet;", sep="")  }
          else                                {  geneSampleValue <- paste("Som;", sep="")  }
        }
      }
    else if (nrow(geneSomaticDataFrame) <= 0)
      {
        if(nrow(geneGermlineDataFrame) > 0)
        {
          if(nrow(geneMethylationData) > 0 )  {  geneSampleValue <- paste("GerMet;", sep="")  }
          else                                {  geneSampleValue <- paste("Ger;", sep="")  }
        }
      }
      
    geneValues <- c(geneValues, geneSampleValue)
  }
  cMat <- rbind(cMat, geneValues)
}

rownames(cMat) <- uniqueSomaticGenes
colnames(cMat) <- uniqueCombinedSamples
```


##Create heatmap for combined data (with methylation)

```{r}
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#FFFFFF", col = NA))
    },
    ##GerSomMet = function(x, y, w, h) {
    ##    grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#C6220F", col = NA))
    ##},
    GerSom = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#4AE4DE", col = NA))
    },
    SomMet = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#FE8514", col = NA))
    },
    Som = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#101DBB", col = NA))
    },
    GerMet = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#E34AE4", col = NA))
    },
    Ger = function(x, y, w, h) {
        grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#39D83F", col = NA))
    }
    
)

col = c("GerSomMet" = "#C6220F", "GerSom" = "#4AE4DE", "SomMet" = "#FE8514", "Som" = "#101DBB", "GerMet" = "#E34AE4", "Ger" = "#39D83F") 

mutationOptions <- c("GerSomMet", "GerSom", "SomMet", "Som", "GerMet", "Ger")

oncoPrint(cMat, get_type = function(x) strsplit(x, ";")[[1]],
    alter_fun = alter_fun, col = col, 
    column_title = "Combined Mutations",
    heatmap_legend_param = list(title = "Value", at = mutationOptions,
        labels = mutationOptions))
```

## Testing

```{r}
alter_fun = list(
    background = function(x, y, w, h) {
        grid.rect(x, y, w*1.9, h*0.9, gp = gpar(fill = "#FFFFFF", col = NA))
    },
    ##GerSomMet = function(x, y, w, h) {
    ##    grid.rect(x, y, w-unit(0.5, "mm"), h-unit(0.5, "mm"), gp = gpar(fill = "#C6220F", col = NA))
    ##},
    GerSom = function(x, y, w, h) {
        grid.rect(x, y, w*1.9, h*0.9, gp = gpar(fill = "#4AE4DE", col = NA))
    },
    SomMet = function(x, y, w, h) {
        grid.rect(x, y, w*1.9, h*0.9, gp = gpar(fill = "#FE8514", col = NA))
    },
    Som = function(x, y, w, h) {
        grid.rect(x, y, w*1.9, h*0.9, gp = gpar(fill = "#101DBB", col = NA))
    },
    GerMet = function(x, y, w, h) {
        grid.rect(x, y, w*1.9, h*0.9, gp = gpar(fill = "#E34AE4", col = NA))
    },
    Ger = function(x, y, w, h) {
        grid.rect(x, y, w*1.9, h*0.9, gp = gpar(fill = "#39D83F", col = NA))
    }
    
)

col = c("GerSomMet" = "#C6220F", "GerSom" = "#4AE4DE", "SomMet" = "#FE8514", "Som" = "#101DBB", "GerMet" = "#E34AE4", "Ger" = "#39D83F") 

mutationOptions <- c("GerSomMet", "GerSom", "SomMet", "Som", "GerMet", "Ger")

oncoPrint(cMat, get_type = function(x) strsplit(x, ";")[[1]],
    alter_fun = alter_fun, col = col, 
    column_title = "Combined Mutations",
    heatmap_legend_param = list(title = "Value", at = mutationOptions,
        labels = mutationOptions))
```
