---
title: "Rtsne_gene_expression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

```{r}
library(readr)
library(Rtsne)
library(dplyr)
library(ggplot2)
library(reshape)
```

## Import the necessary datasets into a dataframe

```{r}
#gene_expression_data <- (read_tsv("../Data/brca_pancan.txt"))
#sampleIDs <- as.vector(as.data.frame(gene_expression_data)[,1])
```

## Import HiSeq data 

```{r}
hiseq_data <- read_tsv("../Data/brca_hiseq.txt")
hiseq_data <- hiseq_data[,-ncol(hiseq_data)] # Remove last column because it has blank values
hiseq_data <- select(hiseq_data, Sample_ID, ERBB2)

#hiseq_data <- as.data.frame(read_tsv("../Data/brca_hiseq.txt"))
#hiseq_IDs <- as.vector(hiseq_data)[,1]
#hiseq_data <- hiseq_data[,-1]
#hiseq_data <- hiseq_data[,-ncol(hiseq_data)] # Remove last column because it has blank values
#rownames(hiseq_data) <- hiseq_IDs
```

## Import GMS data 

```{r}
#gms_feat_data <- (read_tsv("../Data/gms_feature_ERBB2.txt"))
#gms_feat_IDs <- as.vector(as.data.frame(gms_feat_data)[,1])
#gms_fpkm_data <- (read_tsv("../Data/gms_fpkm_ERBB2.txt"))
#gms_fpkm_IDs <- as.vector(as.data.frame(gms_fpkm_data)[,1])

# gms_tpm_data <- as.data.frame(read_tsv("../Data/gms_tpm_ERBB2.txt"))
# gms_tpm_IDs <- as.vector(gms_tpm_data)[,1]
# gms_tpm_data <- as.matrix(gms_tpm_data[,-1,drop=FALSE])
# rownames(gms_tpm_data) <- gms_tpm_IDs

gms_tpm_data <- read_tsv("../Data/gms_tpm_ERBB2.txt")
```

## Load subtype data

```{r}
subtypeData <- read_tsv("../Data/TCGA_Subtypes_Mutations_BasicTabFormatted.txt")
colnames(subtypeData)[1] <- "Sample_ID"
subtypeData <- select(subtypeData, Sample_ID, PAM50)

rowsToKeep <- which(apply(subtypeData, 1, function(x) { !any(is.na(x))}))
subtypeData <- subtypeData[rowsToKeep,]
subtypeData$PAM50 <- factor(subtypeData$PAM50, levels=c("Normal", "LumA", "LumB", "Her2", "Basal"))
```

## PAM50 genes
##  Altered, because one gene wasn't present in the available data
## 1/2-SBSRNA4	A1BG	A1BG-AS1	A1CF	A2LD1	A2M	A2ML1	A2MP1	A4GALT

```{r}
# pam_gene_short <- c("UBE2T", "BIRC5", "NUF2", "CDC6", "CCNB1")
# gms_genes <- c("1/2-SBSRNA4",	"A1BG",	"A1BG-AS1",	"A1CF",	"A2LD1",	"A2M",	"A2ML1",	"A2MP1",	"A4GALT")
# 
# gene_pam <- as.matrix(as.data.frame(subset(gene_expression_data, select=pam_gene_short)))
# row.names(gene_pam) <- sampleIDs
# gene_his <- as.matrix(as.data.frame(subset(hiseq_data, select=pam_gene_short)))
# row.names(gene_his) <- hiseq_IDs
```

## Merge replicate samples

```{r}
shortenSampleIDs <- function(longSampleIDs)
{
  return(as.character(substr(longSampleIDs,1,12)))
}

# shortSamplePam <- shortenSampleIDs(rownames(gene_pam))
# shortSamplePam <- unique(shortSamplePam)
# uniqueShortSamplePam <- unique(shortSamplePam)
# mod_gene_pam <- NULL
# 
# for (sampleID in uniqueShortSamplePam)
# {
#   numReplicates <- table(shortSamplePam)[sampleID]
#   
#   if (numReplicates == 1) {
#     sampleData <- gene_pam[grep(sampleID, rownames(gene_pam)),,drop=FALSE]
#   } else {
#     sampleData <- gene_pam[which(shortenSampleIDs(rownames(gene_pam)) == sampleID),]
#     sampleData <- matrix(apply(sampleData, 2, median), nrow=1)
#     colnames(sampleData) <- colnames(gene_pam)
#   }
# 
#   rownames(sampleData) <- sampleID  
#   mod_gene_pam <- rbind(mod_gene_pam, sampleData)
# }

#print(dim(mod_gene_pam))
```

## Shorten sample names

```{r}
hiseq_data$Sample_ID <- shortenSampleIDs(hiseq_data$Sample_ID)
gms_tpm_data$Sample_ID <- shortenSampleIDs(gms_tpm_data$Sample_ID)

# shortSampleHis <- shortenSampleIDs(rownames(gene_his))
# uniqueShortSampleHis <- unique(shortSampleHis)
# mod_gene_his <- NULL
# 
# for (sampleID in uniqueShortSampleHis)
# {
#   numReplicates <- table(shortSampleHis)[sampleID]
#   
#   if (numReplicates == 1) {
#     sampleData <- gene_his[grep(sampleID, rownames(gene_his)),,drop=FALSE]
#   } else {
#     sampleData <- gene_his[which(shortenSampleIDs(rownames(gene_his)) == sampleID),]
#     sampleData <- matrix(apply(sampleData, 2, median), nrow=1)
#     colnames(sampleData) <- colnames(gene_his)
#   }
# 
#   rownames(sampleData) <- sampleID  
#   mod_gene_his <- rbind(mod_gene_his, sampleData)
# }

#print(dim(mod_gene_his))
```

## Shorten more sample IDs

```{r}
# row.names(gene_his_alt) <- shortenSampleIDs(rownames(gene_his_alt))
# row.names(gms_feat) <- shortenSampleIDs(rownames(gms_feat))
# row.names(gms_fpkm) <- shortenSampleIDs(rownames(gms_fpkm))
# row.names(gms_tpm) <- shortenSampleIDs(rownames(gms_tpm))
```

## Merge Hi-Seq and Pancan datasets

```{r}
colnames(hiseq_data)[2] <- "ERBB2_HiSeq"
colnames(gms_tpm_data)[2] <- "ERBB2_GMS"

merged_data_erbb2 <- inner_join(hiseq_data, gms_tpm_data)
merged_data_erbb2$ERBB2_GMS <- log2(merged_data_erbb2$ERBB2_GMS + 1)

merged_data_erbb2 <- inner_join(merged_data_erbb2, subtypeData)

# merge_request <- function(gene_name) {
#   
#   temp_df <- NULL
#   for(sample_y in shortSamplePam)
#   {
#     combo_zest <- NULL
#     combo_zest <- c(as.numeric(mod_gene_his[sample_y,gene_name]), log2(1 + as.numeric(mod_gene_pam[sample_y,gene_name])))
#     temp_df <- rbind(temp_df, combo_zest)
#   }
#   colnames(temp_df) <- c("HiSeq", "Pancan")
#   
#   return(as.data.frame(temp_df))
# }
```

## Get DFs for genes UBE2T  BIRC5   NUF2    CDC6   CCNB1

```{r}
# ubet <- merge_request("UBE2T")
# birc <- merge_request("BIRC5")
# nuf2 <- merge_request("NUF2")
# cdc6 <- merge_request("CDC6")
# ccnb <- merge_request("CCNB1")

```

## Plot the data

```{r}
# c <- ggplot(sub_feat, aes(x=as.numeric(sub_feat$DataSet1), y=as.numeric(sub_feat$DataSet2), color=sub_feat$PAM50)) + scale_color_manual(name="PAM50", values = c("Normal" = "yellow", "LumA" = "black", "LumB" = "orange", "Her2" = "blue", "Basal" = "Green"))
# c + geom_point() + xlab("HiSeq") + ylab("Feature_counts")
# ```
# 
# ```{r}
# c <- ggplot(sub_fpkm, aes(x=as.numeric(sub_fpkm$DataSet1), y=as.numeric(sub_fpkm$DataSet2), color=sub_fpkm$PAM50)) + scale_color_manual(name="PAM50", values = c("Normal" = "yellow", "LumA" = "black", "LumB" = "orange", "Her2" = "blue", "Basal" = "Green"))
# c + geom_point() + xlab("HiSeq") + ylab("FPKM")
# ```
# 
# ```{r}
# c <- ggplot(sub_tpm, aes(x=as.numeric(sub_tpm$DataSet1), y=as.numeric(sub_tpm$DataSet2), color=sub_tpm$PAM50)) + scale_color_manual(name="PAM50", values = c("Normal" = "yellow", "LumA" = "black", "LumB" = "orange", "Her2" = "blue", "Basal" = "Green"))
# c + geom_point() + xlab("HiSeq") + ylab("TPM")



c <- ggplot(merged_data_erbb2, aes(x=ERBB2_HiSeq, y=ERBB2_GMS, color=PAM50))
c + geom_point() + xlab("HiSeq") + ylab("TPM")

ggsave(filename="../Thesis_images/HisPan_compare.png", height = 8, width = 9, units = "in")
```


