---
title: "Rtsne_mut_sigs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

```{r}
library(readr)
library(Rtsne)
library(dplyr)
library(ggplot2)
```

## Import the necessary datasets into a dataframe

```{r}

mut_sigs_origin <- (read_tsv("../Data/aggregate_weights.txt"))
sampleIDs <- as.vector(as.data.frame(mut_sigs_origin)[,1])

```

## Format the data (be carefull not to run this block more than once, as it will cut off data rows)

```{r}

mut_sigs <- mut_sigs_origin[,-1]
mut_sigs <- as.matrix(mut_sigs)
mut_sig_rows <- head(sampleIDs,-27)
mut_sig_rows <- substr(mut_sig_rows, 1, nchar(mut_sig_rows)-3)
mut_sigs <- head(mut_sigs, -27)

# mut_sig_rows <- unique(mut_sig_rows)
# mut_sigs <- unique(mut_sigs)

row.names(mut_sigs) <- mut_sig_rows
mut_sig_cols <- colnames(mut_sigs)


mut_sigs_2 <- unique(mut_sigs)
```

## Remove the duplicate samples (problematic later)

```{r}

found_samples <- unique(mut_sig_rows)
test_lemon <- seq(1, 981, by=1)

other_df <- NULL

sketchy_rows <- NULL

for (i in test_lemon)
{
  if (mut_sig_rows[i] %in% other_df)
  {
    sketchy_rows <- c(sketchy_rows, i)
  }
  else
  {
    other_df <- c(other_df, mut_sig_rows[i]) 
  }
}

mut_sigs_removed <- mut_sigs[-sketchy_rows, ]

```

## Run Rtsne on the imported data set, and plot it

```{r}
rtsne_out <- Rtsne(mut_sigs, check_duplicates = FALSE, verbose = TRUE)
plot(rtsne_out$Y)

```

## Reformat mut_sigs data into tidy format and save to file. (unrelated to Rtsne)

```{r}

mut_sigs_tidy <- mut_sigs

## to do:
## save rownames to DF
## iterate through each element of rowname_DF,
## write new row of rowname_DF[i], signature, weight
## save to file



```

## This section is for getting BRCA1/2 data

```{r}
germlineData <- as.data.frame(read_tsv("../Data/germline_snv_indel.tsv"))
somaticData <- as.data.frame(read_tsv("../Data/BRCA_parsed_hg38.txt"))
```

## Germline - Build matrices

```{r}
uniqueGermlineGenes <- sort(unique(germlineData$Gene))
uniqueSomaticGenes <- sort(unique(somaticData$Gene))
uniqueSamples <- sort(unique(germlineData$Sample_ID))
uniqueSomaticSamples <- sort(unique(somaticData$Sample)) ##Added

uniqueTotalSamples <- c(uniqueSamples, uniqueSomaticSamples) ##A
uniqueTotalSamples <- sort(unique(uniqueTotalSamples$Sample_ID)) ##A

BRCAGenes <- c("BRCA1", "BRCA2")

```

## Combined data - build matricies

```{r}

combinedSamples <- uniqueSomaticSamples
uniqueCombinedSamples <- unique(combinedSamples)

geneCounter <- 0
```

## Create data matrix for combined data

```{r}

cMat <- NULL
for (sample in mut_sig_rows)
{
  BRCAstatus <- NULL

  gene_one <- "BRCA1"
  gene_two <- "BRCA2"
  
  ##for (gene in BRCAGenes)
  {
    
    geneCounter <- 0
    geneSomaticB1 <- filter(somaticData, Gene==gene_one&Sample==sample)
    geneGermlineB1 <- filter(germlineData, Gene==gene_one&Sample_ID==sample)
    geneSomaticB2 <- filter(somaticData, Gene==gene_two&Sample==sample)
    geneGermlineB2 <- filter(germlineData, Gene==gene_two&Sample_ID==sample)
    geneSampleValue <- ""
  
    if (nrow(geneSomaticB1) > 0 || nrow(geneGermlineB1) > 0) 
      {
        ##geneSampleValue <- paste("BRCA1", sep="") 
        if(nrow(geneSomaticB2) > 0 || nrow(geneGermlineB2) > 0)
        {
          geneSampleValue <- paste("BRCA1 and BRCA2", sep="") 
          print("check 1")
        }
        else  
        {
          geneSampleValue <- paste("BRCA1", sep="")
          print("check 2")
        }
      }
    else if (nrow(geneSomaticB2) > 0 || nrow(geneGermlineB2) > 0)
      {
        geneSampleValue <- paste("BRCA2", sep="")
        print("check 3")
      }
    else { geneSampleValue <- paste("Other", sep="") }  
    
    BRCAstatus <- c(sample, geneSampleValue)
  }
  cMat <- rbind(cMat, BRCAstatus)
}
name_holder <- c("Sample", "BRCA_status")

##rownames(cMat) <- BRCAGenes
colnames(cMat) <- name_holder

```

## plot the data using BRCA labels

```{r}
gMat <- as.data.frame(cMat)

#rtsne_out <- Rtsne(mut_sigs, check_duplicates = FALSE, verbose = TRUE)
#plot(rtsne_out$Y, col=gMat$BRCA_status, pch=20)

```

## format the data in a way that can be used with ggplot2

```{r}

rtsne_gg <- rtsne_out$Y

gene_counter <- 0
rtsne_citrus <- NULL
alist<-c(1:982)

for (item in alist)
{
  gene_counter <- gene_counter + 1
  if (gene_counter <= 981)
  {
    temp_container <- NULL
    temp_container <- c(rtsne_gg[gene_counter,], gMat[gene_counter,]) 
    rtsne_citrus <- rbind(rtsne_citrus, temp_container)
  }
  x_lemon = c("X", "Y", "Sample", "BRCA_status")
  colnames(rtsne_citrus) <- x_lemon
}

#rtsne_gg <- as.data.frame(rtsne_gg)
#rownames(rtsne_gg) <- mut_sig_rows

```

## Plotting the data with ggplot2

```{r}

rtsne_citrine <- as.data.frame(rtsne_citrus)
rownames(rtsne_citrine) <- gMat$Sample

## rtsne_citrine <- data.matrix(rtsne_citrus)

c <- ggplot(rtsne_citrine, aes(x="X",y="Y", color=BRCA_status))
c + geom_point()

```




