---
title: "Rtsne_mut_sigs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

```{r}
library(readr)
library(Rtsne)
library(dplyr)
library(ggplot2)
library(reshape)
```

## Import the necessary datasets into a dataframe

```{r}
mut_sigs_origin <- (read_tsv("../Data/aggregate_weights_4.txt"))
sampleIDs <- as.vector(as.data.frame(mut_sigs_origin)[,1])
```

## Format the data (try not to run this block more than once)

```{r}
mut_sigs <- mut_sigs_origin[,-1]
mut_sigs <- as.matrix(mut_sigs)
mut_sig_rows <- sampleIDs

row.names(mut_sigs) <- mut_sig_rows
mut_sig_cols <- colnames(mut_sigs)

mut_sigs_2 <- unique(mut_sigs)
```

## Run Rtsne on the imported data set, and plot it

```{r}
rtsne_out <- Rtsne(mut_sigs, check_duplicates = FALSE, verbose = TRUE)
plot(rtsne_out$Y)
```

## This section is for getting BRCA1/2 data

```{r}
germlineData <- as.data.frame(read_tsv("../Data/germline_snv_indel.tsv"))
somaticData <- as.data.frame(read_tsv("../Data/BRCA_parsed_hg38.txt"))
```

## Germline - Build matrices

```{r}
uniqueGermlineGenes <- sort(unique(germlineData$Gene))
uniqueSomaticGenes <- sort(unique(somaticData$Gene))
uniqueSamples <- sort(unique(germlineData$Sample_ID))
uniqueSomaticSamples <- sort(unique(somaticData$Sample)) ##Added

uniqueTotalSamples <- c(uniqueSamples, uniqueSomaticSamples) ##A
uniqueTotalSamples <- sort(unique(uniqueTotalSamples$Sample_ID)) ##A

BRCAGenes <- c("BRCA1", "BRCA2")
```

## Create data matrix for combined data

```{r}
geneCounter <- 0

cMat <- NULL
for (sample_s in mut_sig_rows)
{
  BRCAstatus <- NULL
  gene_one <- "BRCA1"
  gene_two <- "BRCA2"
  
  {
    geneCounter <- 0
    geneSomaticB1 <- filter(somaticData, Gene==gene_one&Sample==sample_s)
    geneGermlineB1 <- filter(germlineData, Gene==gene_one&Sample_ID==sample_s)
    geneSomaticB2 <- filter(somaticData, Gene==gene_two&Sample==sample_s)
    geneGermlineB2 <- filter(germlineData, Gene==gene_two&Sample_ID==sample_s)
    geneTest1 <- filter(somaticData, Sample==sample_s)
    geneTest2 <- filter(germlineData, Sample_ID==sample_s)
    geneSampleValue <- ""
  
    if (nrow(geneSomaticB1) > 0 || nrow(geneGermlineB1) > 0) 
      {
        if(nrow(geneSomaticB2) > 0 || nrow(geneGermlineB2) > 0)
        {          geneSampleValue <- paste("BRCA1 and BRCA2", sep="")         }
        else  
        {          geneSampleValue <- paste("BRCA1", sep="")        }
      }
    else if (nrow(geneSomaticB2) > 0 || nrow(geneGermlineB2) > 0)
      {        geneSampleValue <- paste("BRCA2", sep="")      }
    else if (nrow(geneTest1) > 0 || nrow(geneTest2) > 0 )
      { geneSampleValue <- paste("Other", sep="") }  
    else 
      { geneSampleValue <- paste("No_data", sep="") }
    
    if (geneSampleValue != "No_data")
      { BRCAstatus <- c(sample_s, geneSampleValue) }
  }
  if (is.null(BRCAstatus) == FALSE)
    { cMat <- rbind(cMat, BRCAstatus) }
}

name_holder <- c("Sample", "BRCA_status")

colnames(cMat) <- name_holder
```

## format the data in a way that can be used with ggplot2

```{r}
gMat <- as.data.frame(cMat)

rtsne_gg <- rtsne_out$Y

gene_counter <- 0
rtsne_citrus <- NULL
alist<-c(1:1044)

for (item in alist)
{
  gene_counter <- gene_counter + 1
  if (gene_counter <= 1044)
  {
    temp_container <- NULL
    temp_container <- c(rtsne_gg[gene_counter,], gMat[gene_counter,]) 
    rtsne_citrus <- rbind(rtsne_citrus, temp_container)
  }
  x_lemon = c("X", "Y", "Sample", "BRCA_status")
  colnames(rtsne_citrus) <- x_lemon
}
```

## Plotting with ggplot2

```{r}
rownames(rtsne_citrus) <- gMat$Sample
rtsne_citrine <- data.frame(rtsne_citrus)

c <- ggplot(rtsne_citrine, aes(x=as.numeric(rtsne_citrine[,1]),y=as.numeric(rtsne_citrine[,2]) , color=as.character(rtsne_citrine[,4]))) + scale_color_manual(name="BRCA status", values = c("4"="black", "3"="orange", "2"="purple", "1"="green"), breaks=c("1", "2", "3", "4"), labels=c("BRCA1", "Both", "BRCA2", "Other"))

c + geom_point() + xlab("X value") + ylab("Y value")
```

## Define a function here to save "BRCA-like" IDs
## All "BRCA-like" here are any that fall under a particular range in the rtsne plot
##  The range is defined as within x1, x2, y1, and y2
##  >x1, <x2, >y1, <y2
##  Returns an object with "BRCA-like" IDs.

```{r}
selectedSamples <- function(rtsne_DF, x_one, x_two, y_one, y_two)
{ 
  brcaLikeDF <- NULL
  apply(rtsne_DF, 1, function(row)
  {
    sample_container <- NULL
    if (as.numeric(as.character(row[4])) == 4)
    {
      if (as.numeric(row[1]) > x_one && as.numeric(row[1]) < x_two && as.numeric(row[2]) > y_one && as.numeric(row[2]) < y_two)
      {
        sample_container <- c(row.names(row), "BRCA-like", "Mutation_sig")
      }
      else  { sample_container <- c(row.names(row), "Other", "None") }
    }
    else if(as.numeric(row[4] == 1))
    {
      sample_container <- c(row.names(row), "BRCA1", "Mutation")
    }
    else if(as.numeric(row[4] == 2))
    {
      sample_container <- c(row.names(row), "BRCA1&2", "Mutation")
    }
    else if(as.numeric(row[4] == 3))
    {
      sample_container <- c(row.names(row), "BRCA2", "Mutation")
    }
    brcaLikeDF <<- rbind(brcaLikeDF, sample_container)
  } )
  row.names(brcaLikeDF) <- row.names(rtsne_DF)
  colnames(brcaLikeDF) <- c("BRCA_status", "Evidence")
  return(brcaLikeDF)
}
```

## Run the above function, save the output to a file.
##  Note: since the output of rtsne is somewhat randomized, this needs to have
##  X and Y values input manually, I should find a way to automize this eventually
##  X1 and Y1 should be the smaller values, X2 and Y2 should be larger

```{r}
brcaLikeness <- selectedSamples(rtsne_citrus, -2, 12, -30, -14)
```

## Save brcaLikeness DF to file

```{r}
write.table(brcaLikeness, file="brca-likeness_mut_sig.txt", append= FALSE, quote=FALSE, sep="\t", row.names=TRUE, col.names=NA)
```
