---
title: "CancerSubtype"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```

## Read in data

```{r}
TCGASubtypeData <- (read_tsv("../Data/TCGA_Subtypes_Mutations_BasicTabFormatted.txt"))
TCGASubtypeData <- TCGASubtypeData[,colSums(is.na(TCGASubtypeData))<nrow(TCGASubtypeData)]
```

## HRR Gene list

```{r}
#HRR_genes <- c("BRCA1", "BRCA2", "RAD51", "RAD51B", "RAD51D", "DMC1", "XRCC2", "XRCC3", "RAD52", "RAD54L", "RAD54B", "SHFM1", "RAD50", "MRE11A", "NBN", "RBBP8", "MUS81", "EME1", "EME2", "GIYD1", "GIYD2", "GEN1")

## use this list for a less confusing chart
HRR_genes <- c("BRCA1", "BRCA2", "RAD51", "XRCC2", "XRCC3", "RAD52", "RAD54B", "SHFM1", "RAD50", "MRE11A", "NBN", "RBBP8", "EME1", "EME2", "GIYD1", "GIYD2")
```

## Get somatic data

```{r}
#germlineMutationData <- read_tsv("../Data/germline_snv_indel.tsv")

#germlineMutationFrequencies <- table(germlineMutationData$Gene)
#topGermlineMutationGenes <- names(head(sort(germlineMutationFrequencies, decreasing=TRUE), n=8))
somaticMutationData <- as.data.frame(read_tsv("../Data/BRCA_parsed_hg38.txt"))

somaticMutationFrequencies <- table(somaticMutationData$Gene)

somaticMutationPlotData <- filter(somaticMutationData, Gene %in% HRR_genes)
somaticMutationPlotData$Gene <- factor(somaticMutationPlotData$Gene)
```

## Sort by subtype, sort by whether or not there are mutations in BRCA1, etc.

```{r}
subtypeSamples <- sort(unique(TCGASubtypeData$Patient_ID))
subtypeCategory <- unique(sort(TCGASubtypeData$PAM50))
subtypeData <- filter(TCGASubtypeData, PAM50 %in% subtypeCategory)

subtypeDataFrame <- NULL

## inner_join(x, y, by = NULL, copy = FALSE, suffix = c(".x", ".y"), ...)
comboContainer <- left_join(somaticMutationPlotData, subtypeData, by = c("Sample" = "Patient_ID"))

##subtypeGermlinePlotData <- filter(comboContainer, Gene %in% topGermlineMutationGenes)
##subtypeGermlinePlotData$Gene <- factor(subtypeGermlinePlotData$Gene)
```

## Create bar chart for top 10 genes

```{r}
c <- ggplot(comboContainer, aes(x=reorder(PAM50,PAM50, function(x)-length(x)), fill=Gene.x))
c + geom_bar(color = "black") + xlab("Somatic mutations by subtype") ###+ scale_fill_brewer()

#ggsave(filename="../Thesis_images/Subtype_somatic.png", height = 8, width = 9, units = "in")
```

## Create bar chart for all genes

```{r}
##c <- ggplot(comboContainer, aes(x=reorder(PAM50,PAM50, function(x)-length(x)), fill=Gene.x))
##c + geom_bar(color = "black") + xlab("Somatic mutations by subtype") 
```

## Create bar chart for genes of interest 
############# 
## This won't work as is, just modify the code by adding in a gene list in the filter line
## Also, comboContainer is filtered above, create a non-filtered DF to use
#############  

```{r}
GOI_container <- comboContainer
#GOI_container <- filter(GOI_container, Gene.x %in% ###insert gene list of choice###  )
c <- ggplot(GOI_container, aes(x=reorder(PAM50,PAM50, function(x)-length(x)), fill=Gene.x))
c + geom_bar(color = "black") + xlab("Somatic mutations by subtype") ###+ scale_fill_brewer()
```

