---
title: "CancerSubtype"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prep

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(extremevalues)
```

## Read in data

```{r}
TCGASubtypeData <- (read_tsv("../Data/TCGA_Subtypes_Mutations_BasicTabFormatted.txt"))
TCGASubtypeData <- TCGASubtypeData[,colSums(is.na(TCGASubtypeData))<nrow(TCGASubtypeData)]
```

## Somatic gene list
```{r}
germlineMutationData <- read_tsv("../Data/germline_snv_indel.tsv")

germlineMutationFrequencies <- table(germlineMutationData$Gene)
topGermlineMutationGenes <- names(head(sort(germlineMutationFrequencies, decreasing=TRUE), n=8))
```

## HRR Gene list

```{r}
HRR_genes <- c("BRCA1", "BRCA2", "RAD51", "RAD51B", "RAD51D", "DMC1", "XRCC2", "XRCC3", "RAD52", "RAD54L", "RAD54B", "SHFM1", "RAD50", "MRE11A", "NBN", "RBBP8", "MUS81", "EME1", "EME2", "GIYD1", "GIYD2", "GEN1")
```

## Get methylation data

```{r}
methylationData <- read_tsv("../Data/methyl_brca_re.tsv")

methylationProbe <- table(methylationData$Probe)
methylationGenes <- topGermlineMutationGenes
methylationPlotData <- filter(methylationData, Probe %in% methylationGenes)

methylationPlotData$Probe <- factor(methylationPlotData$Probe)
uniqueMethylationSamples <- sort(unique(methylationData$Sample))
```

####
## Try extremevalues to get outliers
## This method works. Sortof. Is in the works.
####

```{r}
geneContainer <- "BRCA1"
methylationGene <- filter(methylationData, Probe %in% geneContainer)
methylationOutliers <- getOutliersI(methylationGene$Value, rho=c(1, 1), FLim=c(0.1, 0.9), distribution="exponential")
outlierPlot(methylationGene$Value, methylationOutliers, mode="qq")

print(methylationOutliers)
```

## Create function to loop through multiple genes, get upper limit from extreme values, save methylation outliers to DF, then use a join function to combine with subtypeData
## To do: save outlier information to a dataframe, inner.join() with subtype data

```{r}
geneLoopList <- NULL
#geneLoopList <- c("BRCA1", "BRCA2", "TP53", "CHEK2", "ATM", "XRCC2")
geneLoopList <- HRR_genes

limitContainer <- NULL
outlierContainer <- NULL

for (gene in geneLoopList)
{
  geneContainer <- gene
  methylationGene <- as.data.frame(filter(methylationData, Probe %in% geneContainer))
  methylationSamples <- methylationGene$Sample
  methylationValues <- methylationGene$Value

  if(gene %in% c("BRCA1"))
  {
    distribution <- "exponential"
  }
  else
  {
    distribution <- "exponential"
  }

  methylationOutliers <- getOutliersI(methylationValues, rho=c(1, 1), FLim=c(0.1, 0.9), distribution=distribution)
  outlierPlot(methylationGene$Value, methylationOutliers, mode="qq")
  outlierSamples <- methylationSamples[methylationOutliers$iRight]
  #print(outlierSamples)
  
  for (sample in outlierSamples)
  {
    tempCont <- c(sample, gene)
    outlierContainer <- rbind(outlierContainer, tempCont)
  }
  
  colnames(outlierContainer) <- c("Sample", "Gene")
  #outlierContainer <- as.data.frame(outlierContainer)
}

outlierContainer <- as.data.frame(outlierContainer)
```

```{r}
cheque_test <- select(subtypeData, Patient_ID, PAM50, Gene)
comboContainer <- inner_join(cheque_test, outlierContainer, by = c("Patient_ID" = "Sample"))
comboContainer$PAM50 <- as.factor(comboContainer$PAM50)
```

## Create bar chart for top 10 genes

```{r}
c <- ggplot(comboContainer, aes(x=reorder(PAM50,PAM50, function(x)-length(x)), fill=Gene.y))
c + geom_bar(color = "black") + xlab("Methylation by subtype") + scale_fill_brewer()

ggsave(filename="../Thesis_images/Subtype_methylation.png", height = 8, width = 9, units = "in")
```

## Save the methylation outliers to a file

```{r}
write.table(outlierContainer, file="methylation_outliers.txt", append= FALSE, quote=FALSE, sep="\t", row.names=TRUE, col.names=NA)
```


